Improvement of Korean Morphological Analysis System \\Through Transformer-Based Re-Ranking

### 2023-11-28 contents/draft.tex
Table 5 앞 장으로 이동 subsection 레이블 변경

    \subsection{Korean Morphological Analysis Corpora}\label{subsec3.1}

@@

    \subsection{Training Example Transformation}\label{subsec3.2}

@@

    \subsection{Lattice Construction and Decoding}\label{subsec3.3}

@@

    \section{Re-ranking Model}\label{sec:re-ranking-model}

    \subsection{Motivation and Background}\label{subsec4.1}

@@

    \subsection{Secondary Path Generation}\label{subsec4.2}

@@

    \subsection{BERT-based Re-ranking}\label{subsec4.3}

@@

    \subsection{Two-stage Re-ranking}\label{subsec4.4}

@@

    In the first step, we re-rank the secondary paths generated using the BERT model, as described in Section ~\ref{subsec4.3}.
    In the second step, we introduced another BERT variant that was optimized for a different set of linguistic features or trained on a different dataset.
    This allowed us to perform a fine-grained re-evaluation, further refine the list, and push more contextually accurate paths to the top.

@@

    \subsection{Setup and Data}\label{subsec5.1}

@@

    We transformed this organized morphological corpus using the training-example transformation process described in Section ~\ref{subsec3.2} to generate samples for training the dictionary-based morphological analysis model.

@@

    \subsection{Evaluation Metrics}\label{subsec5.2}

@@

    \subsection{Basic Performance}\label{subsec5.3}

@@

    \subsection{Re-ranking Performance}\label{subsec5.4}

@@

    Next, we performed the BERT-based re-ranking described in Section ~\ref{sec:re-ranking-model} and compared its performances.
    Three pre-trained language models, KPF-BERT, ETRI-ELECTRA, and ETRI-RoBERTa, which are known to perform well in other Korean language understanding tasks, were used to fine-tune the re-ranking model.
    KPF-BERT received Korean sentences as input, ETRI-ELECTRA received morphologically tagged sentences as input, and ETRI-RoBERTa receives morpheme-separated sentences as input.
    For KPF-BERT and ETRI-RoBERTa, there may be problems with model learning because of the separation of morpheme tags into letter units during the tokenization process when the morpheme analysis results were received as input and re-ranked, hence, the morpheme tags of the mid-level classification unit were added as new tokens, and then training was performed.
    As shown in Figure~\ref{fig:ranking}, for these three types of pre-trained language models, we first performed training with a re-ranking model using only the morphological analysis results and then performed training with a second re-ranking model using the top five morphological analysis results from the first re-ranking results along with the input sentences for other types of pre-trained language models.
    Preliminary tests indicate that using the same type of model or input results in a nearly identical retrained model, with no further improvement in performance.

@@

    \subsection{Comparison to Other Studies}\label{subsec5.5}

@@

    \subsection{Traditional Dictionary-based Approaches}\label{subsec2.1}
    The earliest attempts at Korean morphological analysis relied heavily on rule- and dictionary-based methods~\cite{KwonHC1991}.
    These methodologies employ predefined sets of linguistic rules or large dictionaries to detect morphemes and determine parts of speech.
    One of the most significant advantages of this approach is its deterministic nature, which can lead to high accuracy when the input text closely adheres to the rules or dictionaries used.
    However, it is challenging to scale and update them, particularly with the constant evolution of language and the emergence of new words.
    The dynamic nature of language, particularly in the Internet age, has made the maintenance of comprehensive dictionaries labor-intensive.

    \subsection{Syllable-unit Morphological Analysis}\label{subsec2.2}
    To overcome the limitations of dictionary dependence, syllable-by-syllable morphological analysis has emerged as an alternative~\cite{ShimKS2011, LeeCK2013, LeeCH2016, KimHM2016, KimSW2018, ChoiYS2018, KimHM2019, MinJW2019, SongHJ2019, SongHJ2020, YounJY2021, ShinHJ2023}.
    This method either tags each syllable and then applies a base-form restoration dictionary~\cite{ShimKS2011, LeeCH2016} or tags the syllable with the base form already restored~\cite{YounJY2021}.
    One notable drawback is the challenge of accurately pinpointing morpheme boundaries.
    Furthermore, as the sequences increase in length, the system finds it increasingly challenging to understand long-term contextual data.

    \subsection{Deep Learning and Tokenization}\label{subsec2.3}
    The advent of deep learning has reshaped the Korean morphological analysis landscape.
    High-performance deep learning models in NLP employ tokenization, dividing the text into smaller units and then converting these tokens into vectors as an input to the computational model~\cite{Mikolov2013}.
    The attempt to conduct subword tokenization, incorporating separate morphemes to cater to Korean linguistic characteristics is particularly significant.
    Incorporating accurate morpheme analysis results into tokenization can improve overall analysis performance.

>>>>

    \subsection{Korean Morphological Analysis Corpora}\label{subsec:korean-morphological-analysis-corpora}

@@

    \subsection{Training Example Transformation}\label{subsec:training-example-transformation}

@@

    \subsection{Lattice Construction and Decoding}\label{subsec:lattice-construction-and-decoding}

@@

    \section{Re-ranking Model}\label{sec:reranking-model}

    \subsection{Motivation and Background}\label{subsec:motivation-and-background}

@@

    \subsection{Secondary Path Generation}\label{subsec:secondary-path-generation}

@@

    \subsection{BERT-based Re-ranking}\label{subsec:bert-based-reranking}

@@

    \subsection{Two-stage Re-ranking}\label{subsec:two-stage-reranking}

@@

    In the first step, we re-rank the secondary paths generated using the BERT model, as described in Section ~\ref{subsec:bert-based-reranking}.
    In the second step, we introduced another BERT variant that was optimized for a different set of linguistic features or trained on a different dataset.
    This allowed us to perform a fine-grained re-evaluation, further refine the list, and push more contextually accurate paths to the top.

@@

    \subsection{Setup and Data}\label{subsec:setup-and-data}

@@

    We transformed this organized morphological corpus using the training-example transformation process described in Section ~\ref{subsec:training-example-transformation} to generate samples for training the dictionary-based morphological analysis model.

@@

    \subsection{Evaluation Metrics}\label{subsec:evaluation-metrics}

@@

    \subsection{Basic Performance}\label{subsec:basic-performance}

@@

    \subsection{Re-ranking Performance}\label{subsec:reranking-performance}

@@

    Next, we performed the BERT-based re-ranking described in Section ~\ref{sec:reranking-model} and compared its performances.
    Three pre-trained language models, KPF-BERT, ETRI-ELECTRA, and ETRI-RoBERTa, which are known to perform well in other Korean language understanding tasks, were used to fine-tune the re-ranking model.
    KPF-BERT received Korean sentences as input, ETRI-ELECTRA received morphologically tagged sentences as input, and ETRI-RoBERTa receives morpheme-separated sentences as input.
    For KPF-BERT and ETRI-RoBERTa, there may be problems with model learning because of the separation of morpheme tags into letter units during the tokenization process when the morpheme analysis results were received as input and re-ranked, hence, the morpheme tags of the mid-level classification unit were added as new tokens, and then training was performed.
    As shown in Figure~\ref{fig:ranking}, for these three types of pre-trained language models, we first performed training with a re-ranking model using only the morphological analysis results and then performed training with a second re-ranking model using the top five morphological analysis results from the first re-ranking results along with the input sentences for other types of pre-trained language models.
    Preliminary tests indicate that using the same type of model or input results in a nearly identical retrained model, with no further improvement in performance.

@@

    \subsection{Comparison to Other Studies}\label{subsec:comparison-to-other-studies}

@@

    \subsection{Traditional Dictionary-based Approaches}\label{subsec:traditional-dictionary-based-approaches}
    The earliest attempts at Korean morphological analysis relied heavily on rule- and dictionary-based methods~\cite{KwonHC1991}.
    These methodologies employ predefined sets of linguistic rules or large dictionaries to detect morphemes and determine parts of speech.
    One of the most significant advantages of this approach is its deterministic nature, which can lead to high accuracy when the input text closely adheres to the rules or dictionaries used.
    However, it is challenging to scale and update them, particularly with the constant evolution of language and the emergence of new words.
    The dynamic nature of language, particularly in the Internet age, has made the maintenance of comprehensive dictionaries labor-intensive.

    \subsection{Syllable-unit Morphological Analysis}\label{subsec:syllable-unit-morphological-analysis}
    To overcome the limitations of dictionary dependence, syllable-by-syllable morphological analysis has emerged as an alternative~\cite{ShimKS2011, LeeCK2013, LeeCH2016, KimHM2016, KimSW2018, ChoiYS2018, KimHM2019, MinJW2019, SongHJ2019, SongHJ2020, YounJY2021, ShinHJ2023}.
    This method either tags each syllable and then applies a base-form restoration dictionary~\cite{ShimKS2011, LeeCH2016} or tags the syllable with the base form already restored~\cite{YounJY2021}.
    One notable drawback is the challenge of accurately pinpointing morpheme boundaries.
    Furthermore, as the sequences increase in length, the system finds it increasingly challenging to understand long-term contextual data.

    \subsection{Deep Learning and Tokenization}\label{subsec:deep-learning-and-tokenization}
    The advent of deep learning has reshaped the Korean morphological analysis landscape.
    High-performance deep learning models in NLP employ tokenization, dividing the text into smaller units and then converting these tokens into vectors as an input to the computational model~\cite{Mikolov2013}.
    The attempt to conduct subword tokenization, incorporating separate morphemes to cater to Korean linguistic characteristics is particularly significant.
    Incorporating accurate morpheme analysis results into tokenization can improve overall analysis performance.
